---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
data(df)
data(dict_df)
```

# STATAtransfer

<!-- badges: start -->
<!-- badges: end -->

STATAtransfer provides functions to easily add STATA-style variable and cell labels to entire datasets in R. It also includes tools for quick variable recoding, making it easier to work with labeled data and streamline workflows between STATA and R.

## Installation

You can install the development version of STATAtransfer from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("StringhiniLab/STATAtransfer")
```

## Use

This package consists of three functions. 

### 1. `recode_vars()`. 
Many datasets, such as CLSA, use numeric codes for categorical variables. This can make it harder to label axes when you want to produce many plots or to interpret the data at a glance. 
This function replaces all numeric category values in the data frame with their corresponding labels, making the data easier to read and visualize.
For example, if we want to plot the last three variables of this dataset:
```{r example, warning=FALSE, message=FALSE}
library(STATAtransfer)
library(dplyr)
library(ggplot2)
library(purrr)
head(df)
```
```{r}

# Function for plotting
barplot <- function(var, data = df){
data |> 
count(!!sym(var)) |> 
ggplot(aes(x = !!sym(var), y = n)) +
  geom_col() + 
  coord_flip() +
  theme_minimal()
}

# I create plots for three variables
purrr::map(colnames(df)[4:6], barplot)
```
The category levels of each variable are not possible to recognize. To fix this, you can use `recode_vars` to replace the labels directly, using the dictionary as a reference.

```{r}
# dictionary
head(dict_df)

# recoding the variables
df_recoded <- recode_vars(data = df,
            # specify here other numeric columns you would like to ignore
            ignore_colnames = c(id, age), 
            dictionary = dict_df,
            # dictionary column with the variable names
            var_colname = variable, # dictionary column with the variables
            # dictionary column with the category numbers
            num_colname = level_num,
            # dictionary column with the category labels
            str_colname = level_str)
head(df_recoded)
```
```{r}
purrr::map(data = df_recoded, colnames(df_recoded)[4:6], barplot)
```

### 2. Add labels to save the dataset for STATA


